---
title: "Financial Data Analysis with R"
author: "Yuan Tian"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
---

# Introduction

This document presents several fundamental indicators for analyzing financial data in the U.S. stock market. The present material is licensed under a [Creative Commons Attribution-ShareAlike License 3.0](https://creativecommons.org/licenses/by-sa/3.0/). Some of the material is inspired by [Lei&LoneCapital](https://lonecapital.com/). This work is not possible without Lei's inspiration and support. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require(quantmod)) install.packages('quantmod')
if (!require(BatchGetSymbols)) install.packages('BatchGetSymbols')
library(quantmod)
library(BatchGetSymbols)
library(readr)
library(dplyr)
library(tidyr)
library(kableExtra)
library(ggplot2)
library(purrr)
```

```{r eval=FALSE}
#You can either run the datascraping_v0.0.1.R file first to download, clean, process and save the cleaned data in .csv. 
Source("datascraping_v0.0.1.R")

```

```{r loadCleanBreadth,echo=FALSE, message=TRUE, warning=TRUE}
#set the file working directory
processed_folder <-"./data/processed/"
marketbreadthDSName <-"SP500marketBreadthScores"
file_end <- paste0(Sys.Date(),".csv")
scorefilePath <-paste0(processed_folder,paste0(marketbreadthDSName,file_end))
scores <-read.csv(scorefilePath,check.names=FALSE)
scores$ref.date <-as.Date(scores$ref.date)
```


```{r loadCleanDashboard,echo=FALSE, message=TRUE, warning=TRUE}
#set the file working directory
processed_folder <-"./data/processed/"
marketbreadthDSName <-"MarketDashboardProcessed"
file_end <- paste0(Sys.Date(),".csv")
processdashfilePath <-paste0(processed_folder,paste0(marketbreadthDSName,file_end))
leidashboard <-read.csv(processdashfilePath)
#As.Date
leidashboard$ref.date <-as.Date(leidashboard$ref.date)
#Benchmark VTI
benchmark <- leidashboard[leidashboard$Tickers == "VTI",]%>%select(-X)
colnames(benchmark)<-paste0("BM",colnames(benchmark))
leidashboard<-leidashboard %>%left_join(benchmark,by=c("ref.date"="BMref.date"))%>%
  mutate(ReltoBM = `price.close`/`BMprice.close`)
#keep one month's data for the dashboard
leidashboard<-leidashboard%>%filter(ref.date>(Sys.Date()-30))

marketEndofDay <-leidashboard %>% filter(ref.date == max(ref.date))%>%
  select(watchlist,Sector,name,Tickers,Rel1D,Rel5D,Rel20D,AboveEMA20,AboveEMA60,AboveEMA120,C_Rel_EMA20,EMA20_Rel_EMA60,EMA60_Rel_EMA120,Percent)

watchlistcategory <-unique(marketEndofDay$watchlist)
```

```{r loadCleanMyList,echo=FALSE, message=TRUE, warning=TRUE}
#my own list
processed_folder <-"./data/processed/"
file_end <- paste0(Sys.Date(),".csv")
processmylistfilePath <-paste0(processed_folder,paste0("MyListProcessed",file_end))
mylistfile <-read.csv(processmylistfilePath)
mylistfile$ref.date <- as.Date(mylistfile$ref.date)
#keep one month data
mylistfile<-mylistfile%>%filter(ref.date>(Sys.Date()-30))
mylistfile_endofday <-mylistfile%>%filter(ref.date == max(ref.date))%>%
  select(Tickers,Rel1D,Rel5D,Rel20D,AboveEMA20,AboveEMA60,AboveEMA120,C_Rel_EMA20,EMA20_Rel_EMA60,EMA60_Rel_EMA120)

```


# Market Overview

## SP500 Market Breadth


Companies in SP500 were sorted into 11 sectors according to Global Industry Classification Standards (GICS) based on their primary business activities. The figure shows the percentage of companies that are above 20-day simple moving average (20-day SMA) by each sector. The total breadth score for the SP500 is 1100. 


```{r plotBreadth,echo=FALSE, message=TRUE, warning=TRUE}
#Function: set cell color with green/red and transparency
breadthcolorFill <-function(v,scale = 1){
  #true or false fill.
  color <-NULL
  color = ifelse(v/scale>=0 & v/scale<50,rgb(255,0,0,(50-v/scale)*3,maxColorValue = 255),
                 rgb(0,128,0,(v/scale-50)*3,maxColorValue = 255))
  color
}  

scores %>% select(-"")%>%rename(Reference_Date =ref.date)%>%
kbl(booktabs = T) %>%
  kable_classic_2(full_width = F) %>%column_spec(1,width = "8em")%>%
  column_spec(2,width = "6em",background =sapply(scores$`Communication Services`,breadthcolorFill))%>%
  column_spec(3,width = "6em",background =sapply(scores$`Consumer Discretionary`,breadthcolorFill))%>%
  column_spec(4,width = "6em",background =sapply(scores$`Consumer Staples`,breadthcolorFill))%>%
  column_spec(5,background =sapply(scores$Energy,breadthcolorFill))%>%
  column_spec(6, background =sapply(scores$Financials,breadthcolorFill))%>%
  column_spec(7,width = "6em", background =sapply(scores$`Health Care`,breadthcolorFill))%>%
  column_spec(8, width = "6em",background =sapply(scores$Industrials,breadthcolorFill))%>%
  column_spec(9, width = "6em",background =sapply(scores$`Information Technology`,breadthcolorFill))%>%
  column_spec(10, background =sapply(scores$Materials,breadthcolorFill))%>%
  column_spec(11, background =sapply(scores$`Real Estate`,breadthcolorFill))%>%
  column_spec(12, background =sapply(scores$Utilities,breadthcolorFill))%>%
  column_spec(13, background =sapply(scores$Breadth,breadthcolorFill,scale = 11))
```

## U.S. Market Dashboard

### Broad Market

```{r plot, echo=FALSE, message=TRUE, warning=TRUE}
#Function: set cell color with green/red and transparency
FillTransparentColor <-function(v,maxPercent = 10){
  color <-NULL
  color = ifelse(v<=0,rgb(255,0,0,min(1,abs(v/maxPercent))*250,maxColorValue = 255),
                 rgb(0,128,0,min(1,abs(v/maxPercent))*250,maxColorValue = 255))
  color
}

#Function: set cell color with green/red 
FillSolidRedGreenColor <-function(v){
  color <-NULL
  color = ifelse(v<0 ,rgb(80,0,0,250,maxColorValue = 255),"black")
  color
}

DStoPlot<-function(DS_latestday, DS_Full,watchlist_section){
  tempds <-DS_latestday%>%filter(watchlist == watchlist_section)%>%select(-watchlist,-Sector,-Percent)
  tempfull <-DS_Full%>%filter(watchlist == watchlist_section)
  tempfull$Tickers <- factor(tempfull$Tickers, levels=unique(tempfull$Tickers))
  mpg_list <- split(tempfull$price.close, tempfull$Tickers)
  volumn_list <- split(tempfull$volume,tempfull$Tickers)
  rel_list <- split(tempfull$ReltoBM,tempfull$Tickers) 
  #print(names(rel_list))
  inline_plot <- data.frame(Tickers = names(mpg_list), 
                          OneMonth = "", Volume = "",ReltoBM ="")
  inline_plot <-tempds%>%
  left_join(inline_plot,by = "Tickers")
  inline_plot$OneMonth <-""
  inline_plot$Volume <-""
  inline_plot$ReltoBM<-""
  #print(inline_plot$Tickers)
  inline_plot%>%
  kbl(booktabs = TRUE,
      col.names = c(watchlist_section,"Ticker","1D%","5D%","20D%",">EMA20",">EMA60",">EMA120","C/S%","S/M%","M/L%","1M","Volume","Rel BM")) %>%
  kable_classic_2(full_width = T) %>%
  column_spec(12, image = spec_plot(mpg_list, same_lim = FALSE)) %>%
  column_spec(13, image = spec_plot(volumn_list, polymin = 20,same_lim = FALSE))%>%
  column_spec(14, image = spec_plot(rel_list, same_lim = FALSE))%>%
  row_spec(0, bold = T, color = "white", background = "black")%>%
  column_spec(3,color=sapply(inline_plot$Rel1D,FillSolidRedGreenColor),background =sapply(inline_plot$Rel1D,FillTransparentColor))%>%
  column_spec(4,color=sapply(inline_plot$Rel5D,FillSolidRedGreenColor),background =sapply(inline_plot$Rel5D,FillTransparentColor))%>%
  column_spec(5,color=sapply(inline_plot$Rel20D,FillSolidRedGreenColor),background =sapply(inline_plot$Rel20D,FillTransparentColor))%>%
  column_spec(6,color=sapply(inline_plot$AboveEMA20,FillSolidRedGreenColor),background =ifelse(inline_plot$AboveEMA20>0,"green","red"))%>%
  column_spec(7,color=sapply(inline_plot$AboveEMA60,FillSolidRedGreenColor),background =ifelse(inline_plot$AboveEMA60>0,"green","red"))%>%
  column_spec(8,color=sapply(inline_plot$AboveEMA120,FillSolidRedGreenColor),background =ifelse(inline_plot$AboveEMA120>0,"green","red"))%>%
  column_spec(9,color=sapply(inline_plot$C_Rel_EMA20,FillSolidRedGreenColor),background =sapply(inline_plot$C_Rel_EMA20,FillTransparentColor))%>%
  column_spec(10,color=sapply(inline_plot$EMA20_Rel_EMA60,FillSolidRedGreenColor),background =sapply(inline_plot$EMA20_Rel_EMA60,FillTransparentColor))%>%
  column_spec(11,color=sapply(inline_plot$EMA60_Rel_EMA120,FillSolidRedGreenColor),background =sapply(inline_plot$EMA60_Rel_EMA120,FillTransparentColor))%>%
  add_footnote(c(paste0("Date as Market Close:",max(DS_Full$ref.date)),
                 "Rel BM: 1 month relative to benchmark; C:current price; S(hort): EMA20; M(edium): EMA60; L(ong): EMA120"))
}

DStoPlot(marketEndofDay,leidashboard,watchlistcategory[1])
DStoPlot(marketEndofDay,leidashboard,watchlistcategory[2])
DStoPlot(marketEndofDay,leidashboard,watchlistcategory[3])
DStoPlot(marketEndofDay,leidashboard,watchlistcategory[4])

```

### Market by Industry 

```{r plotindustry, echo=FALSE, message=TRUE, warning=TRUE}
#plot industry togther with the breadth.
  tempds <-marketEndofDay%>%filter(watchlist == watchlistcategory[5])%>%select(-watchlist)
  tempfull <-leidashboard%>%filter(watchlist == watchlistcategory[5])
  tempfull$Tickers <- factor(tempfull$Tickers, levels=unique(tempfull$Tickers))
  mpg_list <- split(tempfull$price.close, tempfull$Tickers)
  volumn_list <- split(tempfull$volume,tempfull$Tickers)
  rel_list <- split(tempfull$ReltoBM,tempfull$Tickers) 
  bread_list<-split(tempfull$Percent,tempfull$Tickers)
  #print(names(rel_list))
  inline_plot <- data.frame(Tickers = names(mpg_list), 
                          OneMonth = "", Volume = "",ReltoBM ="",MarketBreadth = "")
  inline_plot <-tempds%>%
  left_join(inline_plot,by = "Tickers")
  inline_plot$OneMonth <-""
  inline_plot$Volume <-""
  inline_plot$ReltoBM<-""
  inline_plot$MarketBreadth <-""
  #print(inline_plot$Tickers)
  
  start_row_by_sector <-inline_plot%>%mutate(id = row_number())%>%group_by(Sector)%>%mutate(row_id = row_number(Percent))%>%filter(row_id == 1)%>%select(Sector,id)
  
  inline_plot%>%select(-Sector)%>%
  kbl(booktabs = TRUE,
      col.names = c("Name","Ticker","1D%","5D%","20D%",">EMA20",">EMA60",">EMA120","C/S%","S/M%","M/L%","Breadth","1M","1M Volume","Rel BM","1M Breadth")) %>%
  kable_classic_2(full_width = T) %>%
  column_spec(13, image = spec_plot(mpg_list, same_lim = FALSE)) %>%
  column_spec(14, image = spec_plot(volumn_list, polymin = 20,same_lim = FALSE))%>%
  column_spec(15, image = spec_plot(rel_list, same_lim = FALSE))%>%
  column_spec(16, image = spec_plot(bread_list, same_lim = TRUE))%>%
  row_spec(0, bold = T, color = "white", background = "black")%>%
  column_spec(3,color=sapply(inline_plot$Rel1D,FillSolidRedGreenColor),background =sapply(inline_plot$Rel1D,FillTransparentColor))%>%
  column_spec(4,color=sapply(inline_plot$Rel5D,FillSolidRedGreenColor),background =sapply(inline_plot$Rel5D,FillTransparentColor))%>%
  column_spec(5,color=sapply(inline_plot$Rel20D,FillSolidRedGreenColor),background =sapply(inline_plot$Rel20D,FillTransparentColor))%>%
column_spec(6,color=sapply(inline_plot$AboveEMA20,FillSolidRedGreenColor),background =ifelse(inline_plot$AboveEMA20>0,"green","red"))%>%
    column_spec(7,color=sapply(inline_plot$AboveEMA60,FillSolidRedGreenColor),background =ifelse(inline_plot$AboveEMA60>0,"green","red"))%>%
    
  column_spec(8,color=sapply(inline_plot$AboveEMA120,FillSolidRedGreenColor),background =ifelse(inline_plot$AboveEMA120>0,"green","red"))%>%
      column_spec(9,color=sapply(inline_plot$C_Rel_EMA20,FillSolidRedGreenColor),background =sapply(inline_plot$C_Rel_EMA20,FillTransparentColor))%>%
  column_spec(10,color=sapply(inline_plot$EMA20_Rel_EMA60,FillSolidRedGreenColor),background =sapply(inline_plot$EMA20_Rel_EMA60,FillTransparentColor))%>%
  column_spec(11,color=sapply(inline_plot$EMA60_Rel_EMA120,FillSolidRedGreenColor),background =sapply(inline_plot$EMA60_Rel_EMA120,FillTransparentColor))%>%
  column_spec(12,background =sapply(inline_plot$Percent,breadthcolorFill))%>%
  add_footnote(c(paste0("Date as Market Close:",max(tempfull$ref.date)),
                 "Rel BM: 1 month relative to benchmark; C:current price; S(hort): EMA20; M(edium): EMA60; L(ong): EMA120"))%>%
  pack_rows(as.character(start_row_by_sector[1,1]), as.numeric(start_row_by_sector[1,2]), as.numeric(start_row_by_sector[2,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[2,1]), as.numeric(start_row_by_sector[2,2]), as.numeric(start_row_by_sector[3,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[3,1]), as.numeric(start_row_by_sector[3,2]), as.numeric(start_row_by_sector[4,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[4,1]), as.numeric(start_row_by_sector[4,2]), as.numeric(start_row_by_sector[5,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[5,1]), as.numeric(start_row_by_sector[5,2]), as.numeric(start_row_by_sector[6,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[6,1]), as.numeric(start_row_by_sector[6,2]), as.numeric(start_row_by_sector[7,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[7,1]), as.numeric(start_row_by_sector[7,2]), as.numeric(start_row_by_sector[8,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[8,1]), as.numeric(start_row_by_sector[8,2]), as.numeric(start_row_by_sector[9,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[9,1]), as.numeric(start_row_by_sector[9,2]), as.numeric(start_row_by_sector[10,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[10,1]), as.numeric(start_row_by_sector[10,2]), as.numeric(start_row_by_sector[11,2]-1), label_row_css = "background-color: #666; color: #fff;")%>%
  pack_rows(as.character(start_row_by_sector[11,1]), as.numeric(start_row_by_sector[11,2]), as.numeric(start_row_by_sector[11,2]), label_row_css = "background-color: #666; color: #fff;")
```

## My Watchlist

```{r plotmylist,echo=FALSE, message=TRUE, warning=TRUE}

  mylistfile$Tickers <- factor(mylistfile$Tickers, levels=unique(mylistfile$Tickers))
  mpg_list <- split(mylistfile$price.close, mylistfile$Tickers)
  volumn_list <- split(mylistfile$volume,mylistfile$Tickers)
  #print(names(rel_list))
  inline_plot <- data.frame(Tickers = names(mpg_list), 
                          OneMonth = "", Volume = "")
  inline_plot <-mylistfile_endofday%>%
  left_join(inline_plot,by = "Tickers")
  inline_plot$OneMonth <-""
  inline_plot$Volume <-""

  
  inline_plot%>%
  kbl(booktabs = TRUE,
      col.names = c("Ticker","1D%","5D%","20D%",">EMA20",">EMA60",">EMA120","C/S%","S/M%","M/L%","1M","1M Volume")) %>%
  kable_classic_2(full_width = F) %>%
  column_spec(11, image = spec_plot(mpg_list, same_lim = FALSE)) %>%
  column_spec(12, image = spec_plot(volumn_list, polymin = 20,same_lim = FALSE))%>%
  row_spec(0, bold = T, color = "white", background = "black")%>%
  column_spec(2,color=sapply(inline_plot$Rel1D,FillSolidRedGreenColor),background =sapply(inline_plot$Rel1D,FillTransparentColor))%>%
  column_spec(3,color=sapply(inline_plot$Rel5D,FillSolidRedGreenColor),background =sapply(inline_plot$Rel5D,FillTransparentColor))%>%
  column_spec(4,color=sapply(inline_plot$Rel20D,FillSolidRedGreenColor),background =sapply(inline_plot$Rel20D,FillTransparentColor))%>%
    column_spec(5,color=sapply(inline_plot$AboveEMA20,FillSolidRedGreenColor),background =ifelse(inline_plot$AboveEMA20>0,"green","red"))%>%
    column_spec(6,color=sapply(inline_plot$AboveEMA60,FillSolidRedGreenColor),background =ifelse(inline_plot$AboveEMA60>0,"green","red"))%>%
  column_spec(7,color=sapply(inline_plot$AboveEMA120,FillSolidRedGreenColor),background =ifelse(inline_plot$AboveEMA120>0,"green","red"))%>%
      column_spec(8,color=sapply(inline_plot$C_Rel_EMA20,FillSolidRedGreenColor),background =sapply(inline_plot$C_Rel_EMA20,FillTransparentColor))%>%
  column_spec(9,color=sapply(inline_plot$EMA20_Rel_EMA60,FillSolidRedGreenColor),background =sapply(inline_plot$EMA20_Rel_EMA60,FillTransparentColor))%>%
  column_spec(10,color=sapply(inline_plot$EMA60_Rel_EMA120,FillSolidRedGreenColor),background =sapply(inline_plot$EMA60_Rel_EMA120,FillTransparentColor))%>%
  add_footnote(c(paste0("Date as Market Close:",max(tempfull$ref.date)),
                 "C:current price; S(hort): EMA20; M(edium): EMA60; L(ong): EMA120"))

```

